Model Finding for (Co)recursive Functions in SMT

### 2015-05-01 tex/main.tex
More work on prelim

%\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}

@@

A \emph{signature} $\Sigma$ consists of 
a set $\ssorts{\Sigma}$ of type symbols and
a set $\sfuns{\Sigma}$ of function symbols.
We assume that signatures always include a Boolean type $\Bool$ and constants 
$\ltrue$ and $\lfalse$ of that type (respectively, for true and false), 
an (infix) equality predicate $\teq$ of type $\sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$,
and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
and a function $\lite$ of type $\Bool \times \sigma \times \sigma \rightarrow \lite$ for each $\sigma \in \ssorts{\Sigma}$.
We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
We write $\terms( t )$ to denote the set of subterms in $t$.
%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.

A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where 
$\Sigma$ is a signature and  $\Mo$ is a class of $\Sigma$-interpretations,
the \emph{models} of $T$.
A $\Sigma$-formula $\varphi$ is 
\emph{$T$-satisfiable} (resp., \emph{$T$-unsatisfiable}) 
if it is satisfied by some (resp., no) interpretation in $\Mo$.
Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
$\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
For a given signature $\Sigma$
the \emph{theory of equality with uninterpreted functions} \euf
consists of the set of all $\Sigma$-interpretations.
We refer to the sort and function symbols in this theory as \emph{uninterpreted}.

%In this paper, we associate quantified formulas with uninterpreted functions (informally, those whose semantics they define).
In this paper, we consider \emph{annotated quantifed formulas} of the form $\forallf{f} \vec x. \varphi$, 
where $f$ is an uninterpreted function from $\sfuns{\Sigma}$, or the distinguished constant $\fnull$.
When $f$ is not $\fnull$, we will refer to $\forallf{f} \vec x. \varphi$ as a \emph{function definition} (for $f$).
We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
and $\exists \vec x. \varphi$ as shorthand for $\neg \forall \vec x. \neg \varphi$.

\begin{definition}
A formula $\varphi$ \emph{well-defined with respect to functions} $S$ if it is of the form:
\begin{eqnarray} \label{eq:wdf}
(\forallf{f_1} \vec x_1. \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n. \varphi_n) \wedge \psi
\end{eqnarray}
where $f_1, \ldots, f_n$ are distinct, 
$S = \{ f_1, \ldots, f_n \}$,
and $\psi$ contains no function definitions.
\end{definition}

@@

A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
For instance, $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified,
whereas $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not. 
Note that a well-specified function definition need not specify a terminating function, e.g.
$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
\ajr{possibly revise}

>>>>

@@

A \emph{signature} $\Sigma$ consists of 
a set $\ssorts{\Sigma}$ of type symbols and
a set $\sfuns{\Sigma}$ of function symbols.
We assume that signatures always include a Boolean type $\Bool$ and constants 
$\ltrue$ and $\lfalse$ of that type (respectively, for true and false), 
an (infix) equality predicate $\teq$ of type $\tau \times \tau$ for each $\tau \in \ssorts{\Sigma}$,
and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
We write $\terms( t )$ to denote the set of subterms in $t$.
%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.

A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that maps
each $\tau \in \ssorts{\Sigma}$ to a non-empty set $\tau^\I$,
the \emph{domain} of $\tau$ in $\I$;
each $f$ of type $\tau_1 \cdots \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$ 
to a total function $f^\I$ of type $\tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where 
$\Sigma$ is a signature and  $\Mo$ is a class of $\Sigma$-interpretations,
the \emph{models} of $T$.
A $\Sigma$-formula $\varphi$ is 
\emph{$T$-satisfiable} (resp., \emph{$T$-unsatisfiable}) 
if it is satisfied by some (resp., no) interpretation in $\Mo$.
Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
$\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
For a given signature $\Sigma$
the \emph{theory of equality with uninterpreted functions} \euf
consists of the set of all $\Sigma$-interpretations.
We refer to the sort and function symbols in this theory as \emph{uninterpreted}.

%In this paper, we associate quantified formulas with uninterpreted functions (informally, those whose semantics they define).
In this paper, we consider \emph{annotated quantifed formulas} of the form $\forallf{f} \vec x. \varphi$, 
where $f$ is an uninterpreted function from $\sfuns{\Sigma}$, or the distinguished constant $\fnull$.
An annotated quantified formula $\forallf{f} \vec x. \varphi$ is a \emph{function definition (for $f$)} if 
$f$ is a non-$\fnull$ function of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
$\vec x$ is a tuple of $n$ variables of type $\tau_1, \ldots, \tau_n$
and $\forallf{f} \vec x. \varphi$ is eqvuivalent to $\forall \vec x. f( \vec x ) \teq t$ for some $t$.
We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
and $\exists \vec x. \varphi$ as shorthand for $\neg \forall \vec x. \neg \varphi$.

\begin{definition}
A formula $\varphi$ is \emph{well-defined with respect to functions} $\{ f_1, \ldots, f_n \}$ if it is of the form:
\begin{eqnarray} \label{eq:wdf}
(\forallf{f_1} \vec x_1. \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n. \varphi_n) \wedge \psi
\end{eqnarray}
where $f_1, \ldots, f_n$ are not $\fnull$ and distinct, 
and $\psi$ contains no function definitions.
We will commonly call $\psi$ the \emph{conjecture} of $\varphi$.
\end{definition}

@@

A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
For instance, $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified,
whereas $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not. 
Note that a well-specified function definition need not specify a terminating function, e.g.
$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
\ajr{revise}
